### 2018-04-24 summarystats/solanaceaedescriptivev2.Rmd
first organization of data

---
title: "Descriptive analysis of Solanaceae"
author: "Rosana Zenil-Ferguson"
date: "4/20/2018"
output:
  html_document: default
  pdf_document: default
---
```{r,error=FALSE, message=FALSE, warning=FALSE}
library("knitr")
setwd('~/Dropbox/solploidy/basicdata')
library("ape")
library("phytools")
library("treeplyr")

solanaceae.tree<-read.tree("solanaceae2013.tre")

#idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy #selfincomp, ploidybyZF
solanaceae.allrecords<-read.csv("solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")

species<-unique(solanaceae.allrecords$fullname)
total.species<-length(species)# 2278 total species
genera<-unique(solanaceae.allrecords$genus)
total.genera<-length(genera)#97 genera

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

mode.csome<-rep(0,total.species)
min.csome<-rep(0,total.species)
mode.ploidy<-rep(0,total.species)
min.ploidy<-rep(0,total.species)
self.incomp<-rep(0,total.species)
life.hist<-rep(0,total.species)

for (i in 1:total.species){
  index<-which(solanaceae.allrecords$fullname==species[i])
  aux1<- solanaceae.allrecords$csomenumber[index]
  if(all(is.na(aux1))){
    mode.csome[i]<-NA
    min.csome[i]<-NA
  }else{
    mode.csome[i]<-min(Mode(aux1))
    min.csome[i]<-min(aux1,na.rm=TRUE)
  }
  aux2<- solanaceae.allrecords$ploidy[index]
  if(all(is.na(aux2))){
    mode.ploidy[i]<-NA
    min.ploidy[i]<-NA
  }else{
    mode.ploidy[i]<-min(Mode(aux2))
    min.ploidy[i]<-min(aux2,na.rm=TRUE)
  }
  self.incomp[i]<-solanaceae.allrecords$selfincomp[index[1]]
  life.hist[i]<-solanaceae.allrecords$lifehistoryclean[index[1]]
}

simplified.solanaceae<-data.frame(species, min.csome, mode.csome,min.ploidy,mode.ploidy,self.incomp, life.hist)

per.genus.counts<-list()
per.genus.table<-list()
per.genus.countploidy<-list()
per.genus.ploidy<-list()
for(j in 1:total.genera){
  index<-which(solanaceae.allrecords$genus==genera[j])
  per.genus.counts[[j]]<-data.frame(species=solanaceae.allrecords$fullname[index], csome.number=solanaceae.allrecords$csomenumber[index])
  per.genus.table[[j]]<-table(per.genus.counts[[j]]$species, per.genus.counts[[j]]$csome.number)
  per.genus.countploidy[[j]]<-data.frame(species=solanaceae.allrecords$fullname[index], ploidy=solanaceae.allrecords$ploidy[index])
   per.genus.ploidy[[j]]<-table(per.genus.countploidy[[j]]$species, per.genus.countploidy[[j]]$ploidy)
}
names(per.genus.counts)<-genera
names(per.genus.countploidy)<-genera
#Bar plots that show the variance of chromosome numbers per genus

for(j in 1:length(genera)){

  aux1<-colnames(per.genus.table[[j]])
  if(length(aux1)>0){
  #print(per.genus.table[[j]])
  #print(kable(per.genus.table[[j]],format = "markdown"))
  #kable(aux2,format = "markdown")
  barplot(per.genus.table[[j]],main=names(per.genus.counts)[j],border="white")
  }
  aux2<-colnames(per.genus.ploidy[[j]])
  if(length(aux2)>0){
  #print(kable(per.genus.countploidy[[j]],format = "markdown"))
  #kable(aux2,format = "markdown")
  barplot(per.genus.ploidy[[j]],main=names(per.genus.countploidy)[j],border="white")
  }
  }

```