### 2019-03-12 manuscript/main.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

\documentclass[11pt]{article}
\usepackage{evolution}

\begin{document}

%--------------------------------------------------
% Titlepage
%--------------------------------------------------

% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phytologist
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
\begin{center}
    \textsc{Interaction between ploidy, breeding system, and lineage diversification}

\vfill

\noindent 
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
J.~Gordon Burleigh$^{2}$,
%
William A.~Freyman$^{3}$,
%
Boris Igi\'c$^{4}$,
%
Itay Mayrose$^{5}$,
%
and Emma E.~Goldberg$^{3}$
%
\end{center}
%
\noindent$^{1}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
%
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
%
\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
%
\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
%
\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
%
\noindent$^\dagger$\textit{Author for correspondence}.\\

\vfill

\noindent\textit{Running head:} 
Ploidy and Breeding Systems in Solanaceae
%\vfill

\noindent \textit{Keywords:} 
Polyploidy,
Breeding System,
Diversification, SSE models

\vfill

\linenumbers

%--------------------------------------------------
% Abstract
%--------------------------------------------------

\clearpage
\input{abstract}

%--------------------------------------------------
% Main Text
%--------------------------------------------------

\clearpage
\input{introduction}
\input{methods}
\input{results}
\input{discussion}

%--------------------------------------------------
\section{Acknowledgements}
%--------------------------------------------------

This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI).  The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
%BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?

%--------------------------------------------------
% References
%--------------------------------------------------

\clearpage

\setstretch{1}
\bibliography{refs}
\bibliographystyle{evolution} % E: remove URLs
\setstretch{\stretchby}

\end{document}