### 2019-07-15 manuscript/main.tex
- Edits through end of Introduction

\documentclass[11pt]{article}
\usepackage{evolution}

\begin{document}

%--------------------------------------------------
% Titlepage
%--------------------------------------------------

% Original title given to New Phytologist:
% {New comparative approaches for the study of breeding system, polyploidy, and lineage diversification} 
%
\begin{center}
    \textsc{Interaction between Ploidy, Breeding System, and Lineage Diversification}

\vfill

\noindent 
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
J.~Gordon Burleigh$^{2}$,
%
William A.~Freyman$^{3}$,
%
Boris Igi\'c$^{4}$,
%
Itay Mayrose$^{5}$,
%
and Emma E.~Goldberg$^{3}$
%
\end{center}
%
\noindent$^{1}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
%
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
%
\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
%
\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
%
\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
%
\noindent$^\dagger$\textit{Author for correspondence}.\\

\vfill

\noindent\textit{Running head:} 
Ploidy and Breeding Systems in Solanaceae
%\vfill

%B: added three! can have eight
\noindent \textit{Keywords:} 
Polyploidy,
Breeding Systems,
Self-incompatibility,
Diversification, 
Ploidy,
SSE models,
Diploidization.

\vfill

\linenumbers

%--------------------------------------------------
% Abstract
%--------------------------------------------------

\clearpage
\input{abstract}

%--------------------------------------------------
% Main Text
%--------------------------------------------------

\clearpage
\input{introduction}
\input{methods}
\input{results}
\input{discussion}

%--------------------------------------------------
\section{Acknowledgements}
%--------------------------------------------------
Special thanks to Bob Thomson for donating computational resources that made possible testing the convergence of the models. 
We also thank Sergei Tarasov for clarifying lumpability concepts, and Carrie Tribble for troubleshooting some ancestral estimate figures. 
The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Binational Science Foundation BSF-2016550 (to IM).
%B: new:
We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.

%--------------------------------------------------
% References
%--------------------------------------------------

\clearpage

\setstretch{1}
\bibliography{refs}
\bibliographystyle{newphytologist}
\setstretch{\stretchby}

\end{document}